Index measures for oak decline severity using phenotypic descriptors

### 2019-12-15 manuscript/manuscript.Rmd
finished LIME results section

High discrimination was found in the phenotypic descriptors between between the sites using supervised random forest classification (Supplementary  figure #a; margin = `r site_rf_margin$Margin %>% signif(digits = 3)`). 
The descriptors that contributed most to this discrimination were those that were related to trees size such as DBH, total height and crown radius and likely reflect the varied tree age and planting density acrosses the nine sites (Supplementary figure #b). 
After the trees size adjustment was applied to removed inter-site variability but maintain intra-site variability, the random forest margin was substantially reduced to `r site_rf_post_correction_margin$Margin %>% signif(digits = 3)` indicating that the majority of the between site variance had been removed.
This enabled comparisons to be made between the trees at the different sites in the context of decline severity.

@@

The DAI was more complex in descriptor contribution many more descriptors contributing more equally than the PDI (Figure #b).
The overall tree size descriptors live crown ratio, DBH and crown volume contributed almost equally to the DAI index model.
These were followed closely by the stem staining and bleeding descriptors such as bleed prevalence. 
The presence of *A. biguttatus* exit holes had very little contribution to the DAI with exit hole density ranked only `r which(DAI_descriptor_importance$Feature == 'Agrilus exit hole density (m^-2)')` th.

@@

The LIME eas used to identify theImportant descriptors identified model-wide aren't necessarily relevant to localised cases.

- Top descriptors for each case same as top descriptors in overall model importance lists for PDI.
- Decline is associated with a reduction in the size and condition of the crown

- Different descriptors locally compared to overall importance for DAI.
- COD trees have greater crown contact than AOD trees.
- Bleed prevalence not included in any local importances
- Descriptors relevant to context

@@

- reasons for selecting machine learing approach
- use of unsupervised over supervised analysis
- Semi-supervised can be used if unsupervised model is unable to pull out decline severity scale
- For given context, descriptors not contributing to the indexes could be removed
- Concession of model interpretability to predictive capability by using a machine learning approach
- seen as black box but LIME and response surfaces enable interpretation and therefore trust

>>>>

High discrimination was found in the phenotypic descriptors between between the sites using supervised random forest classification (Supplementary  figure #a; margin = `r site_rf_margin$Margin %>% signif(digits = 3)`). 
The descriptors that contributed most to this discrimination were those that were related to trees size such as DBH, total height and crown radius and likely reflect the varied tree age and planting density across the nine sites (Supplementary figure #b). 
After the trees size adjustment was applied to removed inter-site variability but maintain intra-site variability, the random forest margin was substantially reduced to `r site_rf_post_correction_margin$Margin %>% signif(digits = 3)` indicating that the majority of the between site variance had been removed.
This enabled comparisons to be made between the trees at the different sites in the context of decline severity.

@@

The DAI was more complex in descriptor contribution many more descriptors contributing more equally than the PDI (Figure #b).
The overall tree size descriptors live crown ratio, DBH and crown volume contributed almost equally to the DAI index model.
These were followed closely by the stem staining and bleeding descriptors such as bleed prevalence. 
The presence of *A. biguttatus* exit holes had very little contribution to the DAI with exit hole density ranked only `r which(DAI_descriptor_importance$Feature == 'Agrilus exit hole density (m^-2)')`^th^.

@@

LIME identified crown volume, condition transparency and radius as the top four descriptors contributing to the PDI in each of the healthy, moderate and severe decline cases using lime (Table #).
These descriptors were also the same top descriptors in the overall model importance lists for the PDI.
Increasing decline severity was associated with a reduction in the size and condition of the crown.
For instance, declining trees were found to have an estimated crown volume below 841m^3^.
Healthy trees were found to have a crown condition greater than 71.3%, moderately declining trees between 47.6% and 71.3% and severly declining trees a condition below 23.9%.

For the DAI, a number of different descriptors were found to be locally important to the decline type cases compared to the overall model importance such as crown contact (Table #).
This likely reflects the greater number of descriptors contributing to the DAI with different descriptors relevant to particular decline contexts.
The important descriptors identified were mainly associated with tree size such as live crown ratio and lower crown radius.
Bleed prevalence and stem symptom descriptors were not found to be important to any of the decline type cases.
Live crown ratio greater than 61.6% in the severe AOD and moderate COD cases.
COD trees were found to have greater crown contact (greater than 50%) compared to AOD trees.

@@

- reasons for selecting machine learing approach
- use of unsupervised over supervised analysis
- Semi-supervised can be used if unsupervised model is unable to pull out decline severity scale
- For given context, descriptors not contributing to the indexes could be removed
- Concession of model interpretability to predictive capability by using a machine learning approach
- seen as black box but LIME and response surfaces enable interpretation and therefore trust
- LIME can be applied to out of place cases
- Important descriptors identified model-wide aren't necessarily relevant to localised cases.
